MA_Results

### 2023-08-16 Markdown/Manuscript/Child_RMDs/MS_Imp_Guidelines.Rmd
about to start on updating example

### Ethical responsibilities of CHEM developers 
We suggest that three core ethical responsibilities of health economic modellers are to make **socially acceptable** value judgments in planning a CHEM project, to ensure that implementation of a CHEM is **fit for purpose** and to promote **socially beneficial** use of the CHEM over its release lifecycle. There is significant scope for health economists to improve fulfilment of each of these responsibilities.

@@

### CHEM attributes associated with ethical modelling practice
The responsibilities of health economic modellers for the planning, implementation and use of CHEMs are easier to state than to measure.  We have therefore focused our assessment criteria for ethical modelling practice on measurable attributes of CHEMs. As described in Table \@ref(tab:timelygls),  we believe that CHEMs that are **transparent**, **reusable** and **updatable** (TRU) suggest that their creators have met a number of core ethical obligations.  We therefore selected these model attributes to use as the basis for deriving ethical assessment criteria.

@@

Transparency has been identified as a core feature of ethical public health modelling practice [@10.3389/fpubh.2017.00068]. Guidance on transparency in health economic modelling recommended that model code and data should be clearly documented, potentially with different versions for technical and non-technical users [@Eddy2012]. Notably, the same guidelines, published over ten years ago, did not include recommendations on sharing model code and data. However, more recent guidance recommends publicly dissemination of healthcare model artefacts using online repository services [@Erdemir2020]. Repositories such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection. An essential component of quality assuring health economic models is verification - ensuring that calculations are correct and consistent with model specifications [@techver2019]. The extensiveness of verification checks in models implemented as software projects can be reported using the concept of code coverage [@ERICWONG2010188] - the proportion of model code that has been explicitly tested. Tests should ideally combine both unit tests (to verify that small, isolated sections of code produce the correct output when run independently) and acceptance tests (to verify that the correct output is produced when multiple code components are run together to perform tasks that meet core user-requirements [@martin2003agile]). The nature and extent of individual model authorship contributions can become unclear when models are implemented over longer time-frames with a large and changing group of collaborators [@thompson2022escape]. This issue can be addressed by use of online repository services such as GitHub [@github2007], that provide citation tools and can transparently record all individual code contributions to a modelling project over its lifecycle.

Model reusability is the second attribute that we identify as signaling an ethical CHEM implementation. Making a CHEM’s code, data and documentation publicly available is increasingly considered good practice, yet is not sufficient to promote model re-use.  Key concepts in health economic model reuse include generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39]. Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task), is good scientific computing practice [@Wilson_2017] and facilitates model generalisability. Transferability is promoted by computational implementations that store model code and data in distinct files and locations (as opposed to embedding data such as parameter values into source code).  Object-oriented programming approaches include a concept of inheritance [@8717448] that can allow models to be first duplicated and then selectively modified [@ready4oop2022]. Modular computational implementations construct models from multiple independently reusable and replaceable sub-models (modules) [@pan2021modular]. To grant permissions to others to use, test and adapt models and their components, health economists can avail of two broad categories of open source licensing options. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. An alternative approach is to use copyleft licenses [@copyleft2022] that can require content users to distribute any derivative works they create under similar open-source arrangements. 
<!-- Compared to using commercial modelling software, authoring CHEMs in an open-source language like R [@RCORE2022] makes it easier to -->
<!-- Using open-source development platforms and licenses can aid both generalizability and transferability. -->
<!-- The choices that CHEM developers make about model implementation and licensing will also shape who can use a model and for what purposes.  -->
<!-- are relatively easier to selectively modify. -->

Models should be updatable so that they remain valid for longer, evolving as new evidence emerges and the systems being modelled change [@Jenkins2021; @garcia2021cost]. Ensuring that a model is regularly reviewed to identify and implement required improvements is a recommended defense against model validity decay [@calder2018computational]. Sustainable maintenance of open source research software requires both a core development team and an active user community [@info:doi/10.2196/20028]. Online communities can be an efficient means of engaging model users in testing each version of a model, identifying issues and suggesting improvements. Services such as GitHub [@github2007] provide collaborative code development tools [@MERGEL2015464] that help elicit, integrate and reconcile contributions from multiple contributors and to ensure each update is uniquely identifiable and retrievable. It is also important that verification checks are rerun with each model update, a task that can be automated using the software development practice of continuous integration [@CI2017].  The risk of model revisions having unintended consequences for third party users can be mitigated through the use of deprecation conventions [@zhou2016api] that take an informative and staged approach to retiring outdated model code and data.
<!-- Currently, the core development team for a CHEM will be typically be funded to produce a project end-point deliverable whose specifications are well defined early in the project. For more complex and multi-purpose CHEMs, particularly those designed to be incorporated into decision support systems, it may be better for development teams to adopt Agile Software Development [@beck2001manifesto], an approach that has been recommended for complex public health software projects [@10.3389/fpubh.2022.899874]. An Agile model will be less clearly specified in the initial project plan, but will instead continually develop in response to the requirements and feedback of users, who are provided with an initial, simplified working version of the model at the earliest feasible opportunity. -->

###	Assessment criteria
For each attribute that suggests ethical CHEM implementations, we identified assessment criteria that can be met using existing technical solutions and which therefore have the potential to be facilitated by software frameworks.

>>>>

## Ethical responsibilities of CHEM developers 
We suggest that three core ethical responsibilities of health economic modellers are to make **socially acceptable** value judgments when planning a CHEM project, to ensure that implementation of a CHEM is **fit for purpose** and to promote **socially beneficial** use of the CHEM once it is released. There is significant scope for health economists to improve fulfilment of each of these responsibilities.

@@

## CHEM attributes associated with ethical modelling practice
Health economist responsibilities during the planning, implementation and use of CHEMs are easier to state than to measure.  However, aspects of ethical modelling practice may be inferred from measurable attributes of CHEMs. As described in Table \@ref(tab:timelygls),  we believe that the creators of transparent, reusable and updatable (TRU) CHEMs are likely to have fulfilled a number of their ethical obligations.  We therefore selected these model attributes to use as the basis for deriving ethical assessment criteria.

@@

Taking steps to make computational models **transparent** is an important part of ethical public health modelling practice [@10.3389/fpubh.2017.00068]. Guidance on transparency in health economic modelling recommended that model code and data should be clearly documented [@Eddy2012]. Notably, the same guidelines, published over ten years ago, did not recommend sharing model code and data. However, more recent healthcare modelling guidance does recommend publicly dissemination of such artefacts [@Erdemir2020]. Online repository services such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection. Ensuring that calculations are correct and consistent with model specifications is an essential part of CHEM quality assurance [@techver2019]. The extensiveness of such verification checks can be reported using the concept of code coverage [@ERICWONG2010188] - the proportion of model code that has been explicitly tested. Tests should ideally combine both unit tests (to verify that small, isolated sections of code produce the correct output when run independently) and acceptance tests (to verify that the correct output is produced when multiple code components are run together to perform tasks that meet core user-requirements [@martin2003agile]). The nature and extent of individual model authorship contributions can become unclear when models are implemented over longer time-frames with a large and changing group of collaborators [@thompson2022escape]. This issue can be addressed by use of online repository services such as GitHub [@github2007], that provide citation tools and can transparently record all individual code contributions to a modelling project over its lifecycle.

A CHEM that is **reusable** also signals ethical modelling practice. Making a CHEM’s code, data and documentation publicly available is increasingly considered good practice, yet is not sufficient to promote model re-use.  Key concepts in health economic model reuse include generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39]. Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) is good scientific computing practice [@Wilson_2017] and promotes selective reuse. Computational implementations that store model code and data in distinct files and locations (as opposed to embedding data such as parameter values into source code) are easier to selectively modify.  Object-oriented programming approaches incorporate a concept called inheritance [@8717448] that can allow duplicate models to be created and selectively modified. Modular models are constructed from multiple independently reusable and replaceable sub-models (modules) [@pan2021modular]. To grant permissions to others to use, test and adapt models and their components, health economists can avail of two broad categories of open source licensing options. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. An alternative approach is to use copyleft licenses [@copyleft2022] that can require content users to distribute any derivative works they create under similar open-source arrangements. 
<!-- Compared to using commercial modelling software, authoring CHEMs in an open-source language like R [@RCORE2022] makes it easier to -->
<!-- Using open-source development platforms and licenses can aid both generalizability and transferability. -->
<!-- The choices that CHEM developers make about model implementation and licensing will also shape who can use a model and for what purposes.  -->
<!-- are relatively easier to selectively modify. -->

Models should be **updatable** so that they remain valid for longer, evolving as new evidence emerges and the systems being modelled change [@Jenkins2021; @garcia2021cost]. Ensuring that a model is regularly reviewed to identify and implement required improvements is a recommended defense against model validity decay [@calder2018computational]. Sustainable maintenance of open source research software requires both a core development team and an active user community [@info:doi/10.2196/20028]. Online communities can be an efficient means of engaging model users in testing each version of a model, identifying issues and suggesting improvements. Services such as GitHub [@github2007] provide collaborative code development tools [@MERGEL2015464] that help elicit, integrate and reconcile contributions from multiple contributors and to ensure each update is uniquely identifiable and retrievable. It is also important that verification checks are rerun with each model update, a task that can be automated using the software development practice of continuous integration [@CI2017].  The risk of model revisions having unintended consequences for third party users can be mitigated through the use of deprecation conventions [@zhou2016api] that take an informative and staged approach to retiring outdated model code and data.
<!-- Currently, the core development team for a CHEM will be typically be funded to produce a project end-point deliverable whose specifications are well defined early in the project. For more complex and multi-purpose CHEMs, particularly those designed to be incorporated into decision support systems, it may be better for development teams to adopt Agile Software Development [@beck2001manifesto], an approach that has been recommended for complex public health software projects [@10.3389/fpubh.2022.899874]. An Agile model will be less clearly specified in the initial project plan, but will instead continually develop in response to the requirements and feedback of users, who are provided with an initial, simplified working version of the model at the earliest feasible opportunity. -->

##	Assessment criteria
For each attribute that suggests ethical CHEM implementations, we identified assessment criteria that can be met using existing technical solutions and which therefore have the potential to be facilitated by software frameworks.

### 2023-08-16 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd
about to start on updating example

To support the development of CHEMs that meet TRU assessment criteria, we have created a prototype software framework called ready4 that works within the popular open-source programming language R [@RCORE2022]. The ready4 software framework is implemented as a combination of R code libraries and online services and is supported by a documentation website.
<!-- We have designed this framework to enable ethical computational implementations of health economic models that are both open-source (for transparency and reusability) and modular (for reusability and updatability). The framework aims to support the standardization and interoperability that are core enablers of implementing a modular and open source approaches.  -->

### R libraries 
A library in the R language will typically depend on multiple other R libraries, each of which may have different authors. As the number of third-party dependencies of an R library grows so does the fragility of that library (e.g. the library may cease to work as intended due to changes in one of its dependency libraries). To reduce the fragility of our framework we implemented it as multiple R libraries rather than one R library. In total we authored six novel R libraries to implement the ready4 framework, all of which have distinct purposes and dependencies (Table \@ref(tab:cpkgs)).

One framework library provides the **foundation** for implementing CHEMs that are modular and open-source. Modular models are comprised of modules that are able to share inputs and outputs with each other and to be run as independent models [@barros2023empowering]. To be implemented as open-source projects modules authored by different teams must be interoperable. The foundation library facilitates these objectives by defining a template module data structure (using R’s S4 class system) from which all model module data structures will be created and a novel syntax that enable module algorithms to be consistently named. The foundation library also contains tools for retrieving web-based information on model modules, datasets and analysis programs authored with the framework and for partially automating updates to a project documentation website.

Three framework libraries are designed to help standardize workflows for **authoring modules**. The R language supports functional and object-oriented programming paradigms [@10.1214/13-STS452]. Both these paradigms are used in the ready4 framework, each facilitated by a dedicated authoring library. The first module authoring library contains tools for writing functions in a consistent house style and then using the resulting standardised naming conventions to automatically generate basic documentation for each function. The second module authoring library contains tools to help streamline and standardise the authoring of module data structures using the framework's module template and for associating methods (a special type of function) with each module. The second module authoring library also supports the automatic generation of basic documentation for each new module. The third module authoring library provides tools for disseminating themed bundles of model modules as R libraries that are:

@@

-	quality assured (each update triggers GitHub continuous integration workflows, including any acceptance and unit tests created by module library authors).

@@

A library for **authoring datasets** contains tools for supplying model modules with data stored in online repositories (hosted on a Dataverse installation or on GitHub), labelling these datasets and then sharing them via online repositories. A library for **authoring analyses** contains tools to help author programs that combine model modules and datasets to undertake analyses. These analysis programs authored with this library can be either self-documenting (code is easy to understand and integrated with plain English explanations of what it does) or trigger the creation of separate documents (e.g. a scientific manuscript).

@@

###	Online services 
Our framework libraries are designed to be used in conjunction with a number of online services. 
<!-- To facilitate its application to our youth mental health project, we established and configured accounts with these required services (see Availability of data and materials). -->

We created a GitHub organisation (a collection of code repositories) where all code (libraries, programs and sub-routines) that we author with the framework is stored and version controlled. We configured individual repositories in our GitHub organisation that are used for code library projects to use GitHub actions to implement continuous integration. By default, code libraries authored with our framework will use continuous integration to assess compliance with policies specified by the Comprehensive R Archive Network (CRAN) [@CRAN2022]. To track our code coverage, we linked our GitHub organisation to an account we established at codecov [@codecov_2022]. To facilitate the creation and hosting of documentation websites, we enabled GitHub Pages in each repository we used for code library development.

@@

###	Documentation website
We developed a framework documentation website (www.ready4-dev.com) that provides guidance to model developers on how to use and contribute improvements to the ready4 software framework and models developed with it. The documentation website is versioned, which means documentation relating to prior versions of framework software can be archived, retrieved and viewed.

The documentation website was developed using the Hugo framework [@hugo_2023], Docsy theme [@docsy_2023] and Algolia DocSearch [@algoliadocsearch_2023] and is hosted using the Netlify [@netlify_2023] service.  We used functions from our foundation framework library to partially automate website updates relating to available CHEM modules, datasets and analysis programs. We linked our Netlify account to our GitHub organisation so that the project website would automatically update whenever the source code in its GitHub repository was edited.

>>>>

To support the development of CHEMs that meet TRU assessment criteria, we have created a prototype software framework called ready4. The software framework aims to provide a toolkit for: (i) enabling modular implementation of CHEMs; (ii) authoring and documenting CHEM modules; (iii) managing the labelling and transfer of CHEM input and output data; and (iv) authoring reproducible analyses that apply CHEMs to compatible data.

To achieve these goals, we have implemented the software framework as R [@RCORE2022] code libraries that integrate with a number of online services and which are supported by a documentation website.
<!-- We have designed this framework to enable ethical computational implementations of health economic models that are both open-source (for transparency and reusability) and modular (for reusability and updatability). The framework aims to support the standardization and interoperability that are core enablers of implementing a modular and open source approaches.  -->

## R libraries 
A library in the R language will typically depend on multiple other R libraries, each of which may have different authors. As the number of third-party dependencies of an R library grows so does the fragility of that library (e.g. the library may cease to work as intended due to changes in one of its dependency libraries). To reduce the fragility of our framework we implemented it as multiple R libraries rather than one R library. In total we authored six novel R libraries to implement the ready4 framework, all of which have distinct purposes and dependencies (Table \@ref(tab:cpkgs)).

One framework library provides a **foundation for modular CHEM implementations**. In modular model implementations, modules need to be able to share inputs and outputs with each other and to be run as independent models [@barros2023empowering]. To achieve this, the foundation framework library defines a template module (using R’s S4 class system), which can be used to create other modules with a common set of inherited properties. One of these inherited properties is a novel syntax of 15 core commands that enable module algorithms to be consistently named. The foundation library also contains tools for retrieving web-based information on model modules, datasets and analysis programs authored with the framework and for partially automating updates to a project documentation website.

Three framework libraries are designed to help streamline and standardise workflows for **authoring CHEM modules** from the template module. The R language supports functional and object-oriented programming paradigms [@10.1214/13-STS452]. Authoring with each of these paradigms is facilitated by a dedicated CHEM module authoring library. The first module authoring library contains tools for writing functions in a consistent house style and then using the standardised naming conventions of that house style to automatically generate basic documentation for each function. The second module authoring library contains tools to help streamline and standardise the authoring of novel CHEM modules with distinctive properties and for associating novel algorithms with these modules using the framework's syntax. The second module authoring library also supports the automatic generation of basic documentation for each new CHEM module. The third module authoring library provides tools for disseminating themed bundles of CHEM modules as R libraries that are:

@@

-	quality assured (each update triggers continuous integration workflows, including any acceptance and unit tests created by module library authors).

@@

A library for **managing CHEM data** contains tools for supplying CHEM modules with input data ingested from local (i.e. a user's computer) or remote (online repositories) locations, labelling CHEM module datasets and exporting CHEM module data to online repositories. A library for **authoring reproducible analyses** contains tools to help write programs that apply CHEM modules to compatible datasets for the purpose of undertaking health economic analyses. These analysis programs can be either self-documenting (code is integrated with plain English explanations of what it does) or trigger the creation of separate documents (e.g. a scientific manuscript).

@@

##	Online services 
Framework libraries are designed to be used in conjunction with a number of third-party online services that we established and configured accounts with. 
<!-- To facilitate its application to our youth mental health project, we established and configured accounts with these required services (see Availability of data and materials). -->

We created a GitHub organisation (a collection of code repositories) where all code (libraries, programs and sub-routines) that we author with the framework is stored and version controlled. We configured individual repositories in our GitHub organisation that are used for code library projects to use GitHub actions to implement continuous integration. By default, code libraries authored with our framework will use continuous integration to assess compliance with policies specified by the Comprehensive R Archive Network (CRAN) [@CRAN2022]. To track our code coverage, we linked our GitHub organisation to an account we established at codecov [@codecov_2022]. To facilitate the creation and hosting of documentation websites, we enabled GitHub Pages in each repository used for code library development.

@@

##	Documentation website
We developed a framework documentation website (www.ready4-dev.com) that provides guidance to model developers on how to use and contribute improvements to the ready4 software framework and models developed with it. The documentation website is versioned, which means documentation relating to prior versions of framework software can be archived, retrieved and viewed.

The documentation website was developed using the Hugo framework [@hugo_2023], Docsy theme [@docsy_2023] and Algolia DocSearch [@algoliadocsearch_2023] and is hosted using the Netlify [@netlify_2023] service.  We used functions from our foundation framework library to partially automate website updates relating to available CHEM modules, datasets and analysis programs. We linked our Netlify account to our GitHub organisation so that the project website would automatically update whenever its  source code (publicly available in a GitHub repository) was edited.

### 2023-08-16 Markdown/Manuscript/Child_RMDs/MS_Results.Rmd
working way through frameworks
about to start on updating example

## Assessing ethical computational implementations
```{r child="../Child_RMDs/MS_Imp_Guidelines.Rmd", echo=FALSE}
```

## Software framework
```{r child="../Child_RMDs/MS_Imp_Toolkit.Rmd", echo=FALSE}
```

>>>>

# Assessing ethical CHEM implementations
```{r child="../Child_RMDs/MS_Imp_Guidelines.Rmd", echo=FALSE}
```

# Software framework
```{r child="../Child_RMDs/MS_Imp_Toolkit.Rmd", echo=FALSE}
```

### 2023-08-16 Markdown/Manuscript/Child_RMDs/MS_Tables.Rmd
about to start on updating example

```{r echo=FALSE}
if(params$X@tables_in_body_lgl){
  child_docs_ls <- list(one = "../Child_RMDs/MS_Tables_1.RMD",
                        two = "../Child_RMDs/MS_Tables_2.RMD",
                        three = "../Child_RMDs/MS_Tables_3.RMD")
}else{
    child_docs_ls <- list(one = NULL,
                      two = NULL)
}
```

>>>>

```{r echo=FALSE}
if(params$X@tables_in_body_lgl){
  child_docs_ls <- list(one = "../Child_RMDs/MS_Tables_1.Rmd",
                        two = "../Child_RMDs/MS_Tables_2.Rmd",
                        three = "../Child_RMDs/MS_Tables_3.Rmd",
                        four = "../Child_RMDs/MS_Tables_4.Rmd")
}else{
    child_docs_ls <- list(one = NULL,
                          two = NULL,
                          three = NULL,
                          four = NULL)
}
```
